---
title: "Master File"
output: html_notebook
---

```{r}
library(tidyverse)
#for general coding
library(janitor)
#for cleaning
library(lubridate)
#for working with dates
library(readxl)
#for working with excel files
library(fuzzyjoin)
#for working with imperfect strings
library(reshape2)
```

# Questions

For the Taking Responsibility investigation, the San Antonio Express-News had several questions to answer:

How many obstruction charges had occurred since 2009?
How many of these obstruction charges also was associated with an incident where a person had a BAC above 0.15?
How many of these obstruction charges were also associated with a crash?
Our work also led to an important additional question:

How many of any charge labeled "obstruction passageway" involved alcohol and how many did not?
* This question is going to be the most important to answer -- we will need to figure it out first

# Uploading, Checking, & Cleaning

## Uploading the scraped csv files, charges

These are files we scraped from the county clerk's office with information on all codes we pulled for

```{r}
#Misdemeanors
dwi_mis_raw <- read_csv("Data/scraped/bexar-misds-DWI-20191001-20191130.csv") %>%
  mutate(`OFFENSE-CODE` = as.character(`OFFENSE-CODE`)) %>%
  mutate(`CASE-CAUSE-NBR` = as.character(`CASE-CAUSE-NBR`)) %>%
  mutate(`ADDR-HOUSE-NBR` = as.character(`ADDR-HOUSE-NBR`)) %>%
  mutate(`SENTENCE` = as.character(`SENTENCE`)) %>%
  mutate(`SID` = as.character(`SID`)) %>%
  mutate(class = "misdemeanor")

#Felonies
dwi_fel_raw <- read_csv("Data/scraped/bexar-felony-DWI-20191001-20191130.csv") %>%
  mutate(`OFFENSE-CODE` = as.character(`OFFENSE-CODE`)) %>%
  mutate(`CASE-CAUSE-NBR` = as.character(`CASE-CAUSE-NBR`)) %>%
  mutate(`ADDR-HOUSE-NBR` = as.character(`ADDR-HOUSE-NBR`)) %>%
  mutate(`SENTENCE` = as.character(`SENTENCE`)) %>%
  mutate(`SID` = as.character(`SID`)) %>%
  mutate(class = "felony")

#Together
dwi_raw <- dwi_mis_raw %>%
  full_join(dwi_fel_raw)

official_dwi_codes <- read_excel("Data/references_to_import/DWI-related offense codes from excel.xlsx", sheet = 1) %>%
  clean_names() %>%
  mutate(county_offense_code = as.character(county_offense_code))
```

Quick look at what we have

```{r}
dwi_raw_codes <- dwi_raw %>%
  group_by(`OFFENSE-CODE`) %>%
  tally() %>%
  arrange(`OFFENSE-CODE`)

dwi_raw_codes
```

```{r}
head(dwi_raw)
```

Codes we wanted to get, but we don't have

```{r}
check_dwi <- official_dwi_codes %>%
  anti_join(dwi_raw, by = c("county_offense_code" = "OFFENSE-CODE"))

check_dwi
```

## Cleaning the charges csv files

Codes we got, but we don't want to look at

```{r}
clean_dwi_anti <- dwi_raw %>%
  anti_join(official_dwi_codes, by = c("OFFENSE-CODE" = "county_offense_code")) %>%
  group_by(`OFFENSE-DESC`) %>%
  tally()

clean_dwi_anti
```

Okay, that seems okay to me -- let's clean up

```{r}
clean_dwi_pre <- dwi_raw %>%
  inner_join(official_dwi_codes, by = c("OFFENSE-CODE" = "county_offense_code")) %>%
  mutate("BIRTHDATE" = mdy(`BIRTHDATE`)) %>%
  mutate("OFFENSE-DATE" = mdy(`OFFENSE-DATE`)) %>%
  mutate("CUSTODY-DATE" = mdy(`CUSTODY-DATE`)) %>%
  mutate("COMPLAINT-DATE" = ymd(`COMPLAINT-DATE`)) %>%
  mutate("CASE-DATE" = mdy(`CASE-DATE`)) %>%
  mutate("SETTING-DATE" = mdy(`SETTING-DATE`)) %>%
  mutate("G-JURY-DATE" = mdy(`G-JURY-DATE`)) %>%
  mutate("DISPOSITION-DATE" = mdy(`DISPOSITION-DATE`)) %>%
  mutate("JUDGEMENT-DATE" = mdy(`JUDGEMENT-DATE`)) %>%
  mutate("SENTENCE-START-DATE" = mdy(`SENTENCE-START-DATE`)) %>%
  mutate("SENTENCE-END-DATE" = mdy(`SENTENCE-END-DATE`)) %>%
  mutate("POST-JUDICIAL-DATE" = mdy(`POST-JUDICIAL-DATE`)) %>%
  mutate("BOND-DATE" = mdy(`BOND-DATE`)) %>%
  mutate("original_sentence_years" = substr(`ORIGINAL-SENTENCE`, 1, 3)) %>%
  mutate("original_sentence_months" = substr(`ORIGINAL-SENTENCE`, 6, 7)) %>%
  mutate("original_sentence_days" = substr(`ORIGINAL-SENTENCE`, 11, 12)) %>%
  mutate("original_sentence_hours" = substr(`ORIGINAL-SENTENCE`, 16, 18)) %>%
  mutate("offense_date_clean" = ymd(`offense_date_clean`)) %>%
  mutate("FULL-NAME" = str_to_title(`FULL-NAME`)) %>%
  mutate("offense_year" = floor_date(`OFFENSE-DATE`, unit = "year")) %>%
  mutate("judgement_year" = floor_date(`JUDGEMENT-DATE`, unit = "year")) %>%
  ## May 10: adding this in so we can sort by DA
  mutate(da_in_power = case_when(
    judgement_year >= as.Date("2009-01-01") & judgement_year <= as.Date("2014-01-01") ~ "Susan Reed", 
    judgement_year >= as.Date("2015-01-01") & judgement_year <= as.Date("2018-01-01") ~ "Nico Lahood",
    judgement_year >= as.Date("2019-01-01") ~ "Joe Gonzales",
    is.na(judgement_year) & offense_year >= as.Date("2009-01-01") & offense_year <= as.Date("2014-01-01") ~ paste("Susan Reed BY OFFENSE YEAR", offense_year, sep = " "),
    is.na(judgement_year) & offense_year >= as.Date("2015-01-01") & offense_year <= as.Date("2018-01-01") ~ paste("Nico Lahood BY OFFENSE YEAR", offense_year, sep = " "),
    is.na(judgement_year) & offense_year >= as.Date("2019-01-01") ~ paste("Joe Gonzales BY OFFENSE YEAR", offense_year, sep = " "),
  )) %>%
  #select(-judgement_year) %>%
  clean_names() %>%
  select(-x1) %>%
  distinct()


clean_dwi_pre
```

This clean_dwi_pre file includes the "obstruction passageway" cases that are not alcohol-related. We will remove that in a later step

```{r}
clean_dwi_offense_group <- clean_dwi_pre %>%
  group_by(offense_description, reduced_offense_desc) %>%
  tally() 
  #filter(grepl("OBSTRUCT", paste(offense_description, reduced_offense_desc)))

clean_dwi_offense_group
```

```{r}
clean_dwi_offense_group_no_reduce <- clean_dwi_pre %>%
  group_by(offense_description) %>%
  tally() 
  #filter(grepl("OBSTRUCT", paste(offense_description, reduced_offense_desc)))

clean_dwi_offense_group_no_reduce
```

Quick reminder of the maximum and minimum dates in this dataset: 

```{r}
max(clean_dwi_pre$offense_date)
```

```{r}
min(clean_dwi_pre$offense_date)
```

## Uploading the scraped csv files, case details

NOTE: Here is where you would need to have files from scraping second round prep AND Ryan's scrape 

Case details are information we pulled from each person's detailed case page. The ryan-scrape folder shows the code we used to scrape tables from specific URLs we generated

```{r}
scraped_april_1 <- read_csv("Data/my_exports/April17/dwi.csv") %>%
  filter(eventDescription != "Timeout error") %>%
  mutate(sid = as.character(sid)) 

## June 6th update -- some URLs weren't scraped so here's what we found. Still some people aren't on the site. 

scraped_april_2 <- read_csv("Data/scraped/ryan/fixedURLs.csv")%>%
  mutate(sid = as.character(sid)) 

scraped_april_3 <- read_csv("Data/scraped/ryan/timeouts.csv")%>%
  mutate(sid = as.character(sid)) 

scraped_april <- rbind(scraped_april_1, scraped_april_2, scraped_april_3)%>%
  mutate(sid = as.character(sid)) 
```

These people should have obstruction in the title

```{r}
check <- scraped_april %>%
  group_by(offense_description, reduced_offense_desc) %>%
  tally()

check
```

## Cleaning the csv data: scraped files

### Prep to eliminate not-alcohol obstruction passageway

Here, I'm making sure the data processes as I want it to be. Particularly -- that dates are processed as dates. I'm looking at a charge that was "reduced" to something more elevated

```{r}
clean_scraped_prep <- scraped_april %>%
  #mutate(birthDate = mdy(birthDate)) %>%
  select(-birthDate) %>%
  mutate(eventDate = mdy(eventDate)) %>%
  mutate(case_cause_nbr = as.character(case_cause_nbr)) %>%
  filter(is.na(reduced_offense_desc) | reduced_offense_desc != "DRIVING WHILE INTOXICATED") %>%
  filter(is.na(reduced_offense_desc) | reduced_offense_desc != "DWI W/BAC 0.15 OR HIGHER") 
  
  #mutate(case_cause_nbr = if_else(case_cause_nbr == "0", "2019CR5617W", case_cause_nbr))


clean_scraped_prep
```

How many people are in this dataset?

```{r}
clean_scraped_prep_ppl <- clean_scraped_prep %>%
  group_by(sid) %>%
  tally()

clean_scraped_prep_ppl
```

There are 19,714 rows, thus 19,714 people who have been charged with obstruction between Jan. 1, 2009 and Jan. 13, 2022

```{r}
clean_scraped_prep_ppl_case <- clean_scraped_prep %>%
  group_by(case_cause_nbr) %>%
  tally()

clean_scraped_prep_ppl_case
```

There are 19,774 rows, thus 19,774 cases of obstruction between Jan. 1, 2009 and Jan. 13, 2022 

This is a person who must be in here. Let's double check their existence

```{r}
hilario <- clean_scraped_prep %>%
  filter(sid == 1087579)

hilario
```

They exist

### About the scraped (detailed case) data

**Just to review**:

full_name = the person's name

url = the source where the data is from

birthDate = the person's birthday --- I think there is a small problem with this field and it got messed up after the scraping incident, but I'm not worried about it

sex = sex

race = race/ethnicity -- it's a weird column, not reliable, but gives a general idea of the individual's race

offense_description = description of offense

reduced_offense_description = if applicable, the description of the reduced offense

court = court it was in

case_cause_nbr = case number

sid = a unique number for an individual

character_cnt = ignore, filed used to help me generate the url

eventDate = so we are pulling information from the table, located on the url. That table has all of the docket related events. This is the date of the docket-related event

eventDescription = so we are pulling information from the table, located on the url. That table has all of the docket related events. This is the description of the docket-related event

**Data to look for**

clean_scrape_prep = dataset that includes EVERY obstruction charge and the docket lines associated with the person and that charge clean_scrape = the work I do below to only look at DWI info in the obstruction charges

**The Charges**

And we are looking at obstruction-related charges. Here are those specific charges and the number of times they occur (column n)

```{r}
count_charges <- clean_scraped_prep %>%
  select(sid, offense_description, reduced_offense_desc) %>%
  distinct() %>%
  group_by(offense_description, reduced_offense_desc) %>%
  tally()  %>%
  arrange(desc(n))

count_charges
```

# Defining alcohol-related cases

## Looking for DWI events

Now -- I need to figure out what events are the best to look at because we care about how many of these people have a DWI-related docket event So this is where I do a key word search. Basically, give me "DWI", "BAC", "BA C 0.15", "DRIVING WHILE INTOXICATED" BUT don't give me "BACK", "BACH", "NO ALCOHOL", or "WURZBAC" (I know to filter these out through trial and error)

```{r}
dwi_events <- clean_scraped_prep %>%
  filter(grepl("DWI", eventDescription) | grepl("BAC", eventDescription) | grepl("BA C 0.15", eventDescription) | grepl("DRIVING WHILE INTOXICATED", eventDescription)) %>%
  filter(!grepl("BACK", eventDescription) & !grepl("BACH", eventDescription) & !grepl("NO ALCOHOL", eventDescription) & !grepl("WURZBAC", eventDescription)) %>%
  group_by(eventDescription) %>%
  tally() %>%
  arrange(desc(n))

dwi_events
```

I looked through this, and I believe these are the charges we want here as defined in the variables: 

```{r}
official_dwi_events <- c("MB 540409 DRIVING WHILE INTOXICATED", "BA C 0.15 OR HIGHER", "MA 540411 DRIVING WHILE INTOXICATED-2D", "BAC 0.15 OR HIGHER", "MA 540403 DRIVING WHILE INTOXICATED 1S", "540409 DRIVING WHILE INTOXICATED","540409 DRIVING WHILE INTOXICATED", "F3 540412 DRIVING WHILE INTOXICATED-3D", "F3 540413 INTOXICA TION ASSAULT", "DRIVING WHILE INTOXICATED", "540416", "530721", "540414", "540415", "540410")

bac_specific_official_dwi_events <- c("BA C 0.15 OR HIGHER", "BAC 0.15 OR HIGHER")
```

Now, I am looking for those specific docket events in our scraped data

```{r}
clean_scraped <- clean_scraped_prep %>%
  filter(grepl(paste(official_dwi_events, collapse= "|"), eventDescription)) %>%
  filter(is.na(reduced_offense_desc) | reduced_offense_desc != "DRIVING WHILE INTOXICATED")  %>%
  filter(is.na(reduced_offense_desc) | reduced_offense_desc != "DWI W/BAC 0.15 OR HIGHER")

clean_scraped
```

Okay -- so I am looking at the specific docket lines I am interested in, which is helpful -- but I need to focus on the cases. 

Super important to note than the "n" column really doesn't matter. It's a tally of the number of times information in the row appears in all of the docket lines. 

```{r}
check_charges <- clean_scraped %>%
  group_by(offense_description, reduced_offense_desc) %>%
  tally()

check_charges
```

```{r}
check_events <- clean_scraped %>%
  group_by(eventDescription) %>%
  tally()

check_events
```

And here are the case numbers we care about!

```{r}
dwi_case_numbers <- clean_scraped %>%
  distinct(case_cause_nbr) %>%
  mutate(case_cause_nbr = as.character(case_cause_nbr))

dwi_case_numbers
```

And here are the case numbers that are obstruction and are actually obstruction (meaning not alcoholic)!

```{r}
nondwi_obst <- clean_dwi_pre %>%
  anti_join(dwi_case_numbers) %>%
  filter(grepl("OBSTRUC", offense_description)) %>%
  filter(offense_description != "OBSTRUCT HIGHWAY-INTOXICATION")

nondwi_obst
```

```{r}
nondwi_obst_cases <- nondwi_obst %>%
  group_by(case_cause_nbr) %>%
  tally()

nondwi_obst_cases
```

The official dwi cases, no more of the regular, non-alcohol passageway obstruction

```{r}
clean_dwi <- clean_dwi_pre %>%
  anti_join(nondwi_obst, by = c("case_cause_nbr", "offense_desc"))

clean_dwi
```

Can't be too careful. Let's double check. 

```{r}
check_charges <- clean_dwi %>%
  group_by(offense_desc, reduced_offense_desc) %>%
  tally()

check_charges
```

```{r}
ppl <- clean_dwi %>%
  group_by(sid) %>%
  tally()

ppl
```

# Analysis

## Scraped Case Information

### General Analysis, all obstruction codes

Now I want to know how many people are in this dataset. The "n" column is the number of times they appear in my dataset of just "DWI-related" docket events. We can ignore that.

```{r}
people_in_here <- clean_scraped %>%
  group_by(sid) %>%
  tally()

people_in_here
```

Answer: of the 19,714 who had an obstruction charge, about 19,444 had some sort of DWI indication. I know this because there are 19,444 rows

```{r}
people_case_in_here <- clean_scraped %>%
  group_by(sid, case_cause_nbr) %>%
  #group_by(sid) %>%
  tally()

people_case_in_here
```

And of the 19,774 cases, 19,492 of them were related to drinking.

### Passageway/Roadway Analysis

Now I want to look at people who only have this charge and see if they have alcohol-related stuff

So we go to the DWI docket event data, and say -- give me only those with the passageway/roadway charge

Now let's count the people

```{r}
passageway <- clean_scraped %>%
  filter(grepl("OBSTRUCT PASSAGEWAY/ROADWAY", offense_description) | grepl("OBSTRUCT PASSAGEWAY/ROADWAY", reduced_offense_desc)) %>%
  filter(is.na(reduced_offense_desc) | reduced_offense_desc != "OBSTRUCT HIGHWAY-INTOXICATION") %>%
  arrange(sid)

passageway
```

Exporting the docket information for my editor

```{r}
#write_csv(passageway, "Data/my_exports/allie_check_passageway.csv")
```

```{r}
passageway_check_docket <- passageway %>%
  group_by(eventDescription) %>%
  tally()

passageway_check_docket
```

Let me double check I'm looking at the charges I want to look at.

```{r}
passageway_offense_desc <- passageway %>%
  select(sid, offense_description, reduced_offense_desc) %>%
  distinct() %>%
  group_by(offense_description, reduced_offense_desc) %>%
  tally()  %>%
  arrange(desc(n))

passageway_offense_desc
```

And now let's count the people with the charge

```{r}
passageway_ppl <- passageway %>%
  group_by(sid) %>%
  tally() 

passageway_ppl

```

There are 3,562 people that have an indication of a DWI in the dataset — though their charge language doesn't have any indication of intoxication

```{r}
passageway_case <- passageway %>%
  group_by(case_cause_nbr) %>%
  tally()  %>%
  arrange(-n)

passageway_case
```

There are 3,563 cases that have an indication of a DWI in the dataset with 3,562 people

But I know the main criteria we want to look for is a BAC above 0.15 or higher. So how many of these people had a BAC of 0.15 or higher?

All the BAC events:

```{r}
pass_BAC_events <- passageway %>%
  filter(grepl(paste(bac_specific_official_dwi_events, collapse="|)"), eventDescription)) 

pass_BAC_events
```

All the BAC people:

```{r}
people_pass_BAC <- pass_BAC_events %>%
  group_by(sid) %>%
  tally()

people_pass_BAC
```

There are 1,385 people with a BAC of 0.15 or higher.

```{r}
case_pass_BAC <- pass_BAC_events %>%
  group_by(case_cause_nbr) %>%
  tally()

case_pass_BAC
```

There are 1,385 cases with a BAC of 0.15 or higher.

But how many people had this passageway roadway charge?

Let's look at the dataset before we filtered for DWI people

```{r}
count_all_passageway <- clean_scraped_prep %>%
  filter(grepl("OBSTRUCT PASSAGEWAY/ROADWAY", offense_description) | grepl("OBSTRUCT PASSAGEWAY/ROADWAY", reduced_offense_desc)) %>%
  filter(is.na(reduced_offense_desc) | reduced_offense_desc != "OBSTRUCT HIGHWAY-INTOXICATION") %>%
  group_by(sid) %>%
  #group_by(sid, case_cause_nbr) %>%
  tally()
  
count_all_passageway
```

**Final Finding**: Of the 3,837 cases with obstruction of passageway charge between Jan. 2, 2009 and Jan. 13, 2022, 3,562 of them have been associated with a drinking above the legal limit. 1,385 people had a blood alcohol content above 0.15.

### Highway Intox Obstruction

Okay, let's look at people with the obstruction highway-intoxication charge.

First all of the docket events with this charge:

```{r}
highway_intox <- clean_scraped %>%
  filter(grepl("OBSTRUCT HIGHWAY-INTOX", offense_description) | grepl("OBSTRUCT HIGHWAY-INTOX", reduced_offense_desc)) %>%
  filter(is.na(reduced_offense_desc) | reduced_offense_desc != "OBSTRUCT PASSAGEWAY/ROADWAY/WA")

highway_intox
```

Just to prove to myself that I am in fact looking at people with these only these charges, let's double check:

```{r}
highway_intox_offense_desc <- highway_intox %>%
  select(sid, offense_description, reduced_offense_desc) %>%
  distinct() %>%
  group_by(offense_description, reduced_offense_desc) %>%
  tally()  %>%
  arrange(desc(n))

highway_intox_offense_desc
```

Now the number of people with this charge

```{r}
case_in_here_highway_intox <- highway_intox %>%
  group_by(case_cause_nbr) %>%
  #group_by(sid) %>%
  tally()

case_in_here_highway_intox
```

```{r}
people_in_here_highway_intox <- highway_intox %>%
  group_by(sid) %>%
  tally()

people_in_here_highway_intox
```

So, there are 15,929 cases charged with (or had their charge reduced to) "OBSTRUCT HIGHWAY-INTOXICATION." 15,891 people

But I know the main criteria we want to look for is a BAC above 0.15 or higher. So how many of these people had a BAC of 0.15 or higher?

All the BAC events:

```{r}
highway_intox_BAC <- highway_intox %>%
  filter(grepl(paste(bac_specific_official_dwi_events, collapse="|)"), eventDescription)) 

highway_intox_BAC
```

All the BAC cases:

```{r}
case_highway_intox_BAC <- highway_intox_BAC %>%
  group_by(case_cause_nbr) %>%
  tally()

case_highway_intox_BAC
```

3,564 cases of BAC  0.15 in highway intoxication

How many people?

```{r}
people_highway_intox_BAC <- highway_intox_BAC %>%
  group_by(sid) %>%
  tally()

people_highway_intox_BAC
```

So there are 3,564 cases who had a BAC above 0.15 3,558 people

How many people in total had a highway intoxication charge?

```{r}
count_all_highway_intox <- clean_scraped_prep %>%
  filter(grepl("OBSTRUCT HIGHWAY-INTOX", offense_description) | grepl("OBSTRUCT HIGHWAY-INTOX", reduced_offense_desc)) %>%
  filter(is.na(reduced_offense_desc) | reduced_offense_desc != "OBSTRUCT PASSAGEWAY/ROADWAY/WA") %>%
  group_by(case_cause_nbr) %>%
  tally()

  
count_all_highway_intox
```

**Final Finding**: Of the 15,929 cases with an obstruction highway intoxication charge between Jan. 1, 2009 and Jan. 13, 2022, 15,929 of them have been associated with a drinking above the legal limit. 3,564 of the cases had a docket line indicating that their BAC was above 0.15.




## BAC master

```{r}
bac_master <- clean_scraped %>%   
  filter(grepl(paste(bac_specific_official_dwi_events, collapse="|)"), eventDescription)) 


bac_ppl <- bac_master %>%
  group_by(sid) %>%
  tally() %>%
  arrange(desc(n))

bac_ppl
```

Check a random person's docket

```{r}
docket_1099456 <- clean_scraped %>%
  filter(sid == 1099456)

docket_1099456
```

## DWI master

```{r}
dwi_master <- clean_scraped %>%   
  filter(grepl(paste(official_dwi_events, collapse="|"), eventDescription)) 


dwi_ppl <- dwi_master %>%
  group_by(sid) %>%
  tally() %>%
  arrange(desc(n))

dwi_ppl
```

## Dated after Oct. 27, 2020

This is for a specific point in the story

```{r}
oct27 <- clean_scraped %>%
  filter(eventDate > as.Date("2020-10-27")) 
 

oct27
```

```{r}
oct27_ppl <- oct27 %>%
 group_by(case_cause_nbr) %>%
  tally()

oct27_ppl
```

490 people

```{r}
oct27_BAC_events <- oct27 %>%
  filter(grepl(paste(bac_specific_official_dwi_events, collapse="|)"), eventDescription)) 

oct27_BAC_events
```

```{r}
oct27_BAC_ppl <- oct27_BAC_events %>%
 group_by(case_cause_nbr) %>%
  tally()

oct27_BAC_ppl
```



