---
title: "DWI: Scraping Prep"
output: html_notebook
---
Steps for Scraping:

https://search.bexar.org/Case/CaseDetail?st=s&s= (INSERT SID NUMBER) &cs=&ct=&=&full=y&p=1_ (INSERT CASE NUMBER) +++++++++ (INSERT COURT) (INSERT JUDICIAL NUMBER) 00000

Final goals: 

1. Distinguish passageway/roadway -- which ones are DWI related

2. How many DWI-related obstruction charges are associated with BACs over 0.15

3. Find obstruction charges that were reduced from intoxication assault
a. This one we can probably just use the existing dataset, look at those who have two charges in the same day

# Tools

```{r}
library("RSelenium")
library("rvest")
library("tidyverse")
library("httr")
library("jsonlite")
library("janitor")
```


# Same code from master-file.Rmd 

Basically to begin, I need to take the files that I scraped and figure out how to then get the docket information from this website: https://search.bexar.org/Case/CaseDetail?st=s&s=

## Uploading, Checking, & Cleaning

### Uploading the scraped csv files, charges

These are files we scraped from the county clerk's office with information on all codes we pulled for

```{r}
#Misdemeanors
dwi_mis_raw <- read_csv("Data/scraped/bexar-misds-DWI-20191001-20191130.csv") %>%
  mutate(`OFFENSE-CODE` = as.character(`OFFENSE-CODE`)) %>%
  mutate(`CASE-CAUSE-NBR` = as.character(`CASE-CAUSE-NBR`)) %>%
  mutate(`ADDR-HOUSE-NBR` = as.character(`ADDR-HOUSE-NBR`)) %>%
  mutate(`SENTENCE` = as.character(`SENTENCE`)) %>%
  mutate(`SID` = as.character(`SID`)) %>%
  mutate(class = "misdemeanor")

#Felonies
dwi_fel_raw <- read_csv("Data/scraped/bexar-felony-DWI-20191001-20191130.csv") %>%
  mutate(`OFFENSE-CODE` = as.character(`OFFENSE-CODE`)) %>%
  mutate(`CASE-CAUSE-NBR` = as.character(`CASE-CAUSE-NBR`)) %>%
  mutate(`ADDR-HOUSE-NBR` = as.character(`ADDR-HOUSE-NBR`)) %>%
  mutate(`SENTENCE` = as.character(`SENTENCE`)) %>%
  mutate(`SID` = as.character(`SID`)) %>%
  mutate(class = "felony")

#Together
dwi_raw <- dwi_mis_raw %>%
  full_join(dwi_fel_raw)

official_dwi_codes <- read_excel("Data/references_to_import/DWI-related offense codes from excel.xlsx", sheet = 1) %>%
  clean_names() %>%
  mutate(county_offense_code = as.character(county_offense_code))
```

Quick look at what we have

```{r}
dwi_raw_codes <- dwi_raw %>%
  group_by(`OFFENSE-CODE`) %>%
  tally() %>%
  arrange(`OFFENSE-CODE`)

dwi_raw_codes
```

```{r}
head(dwi_raw)
```

Codes we wanted to get, but we don't have

```{r}
check_dwi <- official_dwi_codes %>%
  anti_join(dwi_raw, by = c("county_offense_code" = "OFFENSE-CODE"))

check_dwi
```

## Cleaning the charges csv files

Codes we got, but we don't want to look at

```{r}
clean_dwi_anti <- dwi_raw %>%
  anti_join(official_dwi_codes, by = c("OFFENSE-CODE" = "county_offense_code")) %>%
  group_by(`OFFENSE-DESC`) %>%
  tally()

clean_dwi_anti
```

Okay, that seems okay to me -- let's clean up

```{r}
clean_dwi_pre <- dwi_raw %>%
  inner_join(official_dwi_codes, by = c("OFFENSE-CODE" = "county_offense_code")) %>%
  mutate("BIRTHDATE" = mdy(`BIRTHDATE`)) %>%
  mutate("OFFENSE-DATE" = mdy(`OFFENSE-DATE`)) %>%
  mutate("CUSTODY-DATE" = mdy(`CUSTODY-DATE`)) %>%
  mutate("COMPLAINT-DATE" = ymd(`COMPLAINT-DATE`)) %>%
  mutate("CASE-DATE" = mdy(`CASE-DATE`)) %>%
  mutate("SETTING-DATE" = mdy(`SETTING-DATE`)) %>%
  mutate("G-JURY-DATE" = mdy(`G-JURY-DATE`)) %>%
  mutate("DISPOSITION-DATE" = mdy(`DISPOSITION-DATE`)) %>%
  mutate("JUDGEMENT-DATE" = mdy(`JUDGEMENT-DATE`)) %>%
  mutate("SENTENCE-START-DATE" = mdy(`SENTENCE-START-DATE`)) %>%
  mutate("SENTENCE-END-DATE" = mdy(`SENTENCE-END-DATE`)) %>%
  mutate("POST-JUDICIAL-DATE" = mdy(`POST-JUDICIAL-DATE`)) %>%
  mutate("BOND-DATE" = mdy(`BOND-DATE`)) %>%
  mutate("original_sentence_years" = substr(`ORIGINAL-SENTENCE`, 1, 3)) %>%
  mutate("original_sentence_months" = substr(`ORIGINAL-SENTENCE`, 6, 7)) %>%
  mutate("original_sentence_days" = substr(`ORIGINAL-SENTENCE`, 11, 12)) %>%
  mutate("original_sentence_hours" = substr(`ORIGINAL-SENTENCE`, 16, 18)) %>%
  mutate("offense_date_clean" = ymd(`offense_date_clean`)) %>%
  mutate("FULL-NAME" = str_to_title(`FULL-NAME`)) %>%
  mutate("offense_year" = floor_date(`OFFENSE-DATE`, unit = "year")) %>%
  mutate("judgement_year" = floor_date(`JUDGEMENT-DATE`, unit = "year")) %>%
  ## May 10: adding this in so we can sort by DA
  mutate(da_in_power = case_when(
    judgement_year >= as.Date("2009-01-01") & judgement_year <= as.Date("2014-01-01") ~ "Susan Reed", 
    judgement_year >= as.Date("2015-01-01") & judgement_year <= as.Date("2018-01-01") ~ "Nico Lahood",
    judgement_year >= as.Date("2019-01-01") ~ "Joe Gonzales",
    is.na(judgement_year) & offense_year >= as.Date("2009-01-01") & offense_year <= as.Date("2014-01-01") ~ paste("Susan Reed BY OFFENSE YEAR", offense_year, sep = " "),
    is.na(judgement_year) & offense_year >= as.Date("2015-01-01") & offense_year <= as.Date("2018-01-01") ~ paste("Nico Lahood BY OFFENSE YEAR", offense_year, sep = " "),
    is.na(judgement_year) & offense_year >= as.Date("2019-01-01") ~ paste("Joe Gonzales BY OFFENSE YEAR", offense_year, sep = " "),
  )) %>%
  #select(-judgement_year) %>%
  clean_names() %>%
  select(-x1) %>%
  distinct()


clean_dwi_pre
```

## End Duplicate Code

# URL making

```{r}
info_needed <- clean_dwi_pre %>%
  filter(grepl("BSTRUC", c(offense_description)) | grepl("BSTRUC", c(reduced_offense_desc))) %>%
  select(full_name, birthdate, sex, race, offense_date, offense_description, reduced_offense_desc, "court", starts_with(c("case_cause", "sid", "judic"))) 

info_needed
```

```{r}
count <- info_needed %>%
  group_by(sid) %>%
  tally() %>%
  arrange(desc(n))

count
```

```{r}
url_prep <- info_needed %>%
  mutate(character_cnt = nchar(court)) %>%
    mutate(url = case_when(
    !is.na(case_cause_nbr) & character_cnt == 4 ~ paste("https://search.bexar.org/Case/CaseDetail?st=s&s=", sid, "&cs=&ct=&=&full=y&p=1_", case_cause_nbr, "+++++++++", court, judicial_nbr, "00000", sep = ""),
    !is.na(case_cause_nbr) & character_cnt == 3 ~ paste("https://search.bexar.org/Case/CaseDetail?st=s&s=", sid, "&cs=&ct=&=&full=y&p=1_", case_cause_nbr, "+++++++++", court, "+", judicial_nbr, "00000", sep = ""),
    TRUE ~ ""
    )) %>%
  filter(url != "")%>%
  select(full_name, url, everything())

url_prep
```

# Fixing Errors

This is more of a guessing process, not super formulaic 

```{r}
missing <- read_csv("Data/scraped/ryan/DWI Strays - Strays.csv") %>%
  clean_names() %>%
  select(full_name)

missing
```

```{r}
missing_and_needed_info <- missing %>%
  inner_join(info_needed)

missing_and_needed_info
```

Why is there an additional row -- it likely makes sense but let me check 

```{r}
missing_and_needed_info_check <- missing_and_needed_info %>%
  group_by(full_name) %>%
  tally() %>%
  arrange(desc(n))

missing_and_needed_info_check
```


```{r}
miguel_check <- missing_and_needed_info %>%
  filter(full_name == "Hernandez, Miguel")

miguel_check
```

Okay this is why we check -- looks like the Miguel we care about had a passageway offense so we will filter out for the other Miguel

```{r}
clean_missing_and_needed_info <- missing_and_needed_info %>%
  filter(birthdate != as.Date("1985-07-29")) %>%
  arrange(full_name)

clean_missing_and_needed_info
```


```{r}
write_csv(clean_missing_and_needed_info, "Data/my_exports/May13/fix_missing_urls.csv")
```


## timed out: copy paste discovery and fix

###### Okay let me check the event descriptions here

```{r}
count_all_highway_intox_anti_prep <- count_all_highway_intox %>%
  select(-n)

people_in_here_highway_intox_anti_prep <- people_in_here_highway_intox %>%
  select(-n)

highway_anti <- count_all_highway_intox_anti_prep %>%
  anti_join(people_in_here_highway_intox_anti_prep)  %>%
  inner_join(clean_scraped_prep) %>%
  #filter(is.na(reduced_offense_desc))
  group_by(offense_description, reduced_offense_desc) %>%
  tally() 
  #arrange(desc(n)) 

highway_anti
```

### Potentially more missing people

```{r}
clean_scraped_prep_missing <- clean_scraped_prep %>%
  filter(eventDescription == "Timeout error")

clean_scraped_prep_missing
```

```{r}
write_csv(clean_scraped_prep_missing, "code/Data/for_ryan_still_missing.csv")
```

