---
title: "DWI: Scraping Prep"
output: html_notebook
---
Steps for Scraping:

https://search.bexar.org/Case/CaseDetail?st=s&s= (INSERT SID NUMBER) &cs=&ct=&=&full=y&p=1_ (INSERT CASE NUMBER) +++++++++ (INSERT COURT) (INSERT JUDICIAL NUMBER) 00000

1. Distinguish passageway/roadway -- which ones are DWI related

2. How many DWI-related obstruction charges are associated with BACs over 0.15

3. Find obstruction charges that were reduced from intoxication assault
a. This one we can probably just use the existing dataset, look at those who have two charges in the same day


# Original

```{r}
library("RSelenium")
library("rvest")
library("tidyverse")
library("httr")
library("jsonlite")
library("janitor")
```


```{r}
info_needed <- clean_dwi_pre %>%
  filter(grepl("BSTRUC", c(offense_description)) | grepl("BSTRUC", c(reduced_offense_desc))) %>%
  select(full_name, birthdate, sex, race, offense_date, offense_description, reduced_offense_desc, "court", starts_with(c("case_cause", "sid", "judic"))) 

info_needed
```

```{r}
count <- info_needed %>%
  group_by(sid) %>%
  tally() %>%
  arrange(desc(n))

count
```

```{r}
url_prep <- info_needed %>%
  mutate(character_cnt = nchar(court)) %>%
    mutate(url = case_when(
    !is.na(case_cause_nbr) & character_cnt == 4 ~ paste("https://search.bexar.org/Case/CaseDetail?st=s&s=", sid, "&cs=&ct=&=&full=y&p=1_", case_cause_nbr, "+++++++++", court, judicial_nbr, "00000", sep = ""),
    !is.na(case_cause_nbr) & character_cnt == 3 ~ paste("https://search.bexar.org/Case/CaseDetail?st=s&s=", sid, "&cs=&ct=&=&full=y&p=1_", case_cause_nbr, "+++++++++", court, "+", judicial_nbr, "00000", sep = ""),
    TRUE ~ ""
    )) %>%
  filter(url != "")%>%
  select(full_name, url, everything())

url_prep
```

# Fixing Errors

```{r}
missing <- read_csv("Data/scraped/ryan/DWI Strays - Strays.csv") %>%
  clean_names() %>%
  select(full_name)

missing
```

```{r}
missing_and_needed_info <- missing %>%
  inner_join(info_needed)

missing_and_needed_info
```

Why is there an additional row -- it likely makes sense but let me check 

```{r}
missing_and_needed_info_check <- missing_and_needed_info %>%
  group_by(full_name) %>%
  tally() %>%
  arrange(desc(n))

missing_and_needed_info_check
```


```{r}
miguel_check <- missing_and_needed_info %>%
  filter(full_name == "Hernandez, Miguel")

miguel_check
```

Okay this is why we check -- looks like the Miguel we care about had a passageway offense so we will filter out for the other Miguel

```{r}
clean_missing_and_needed_info <- missing_and_needed_info %>%
  filter(birthdate != as.Date("1985-07-29")) %>%
  arrange(full_name)

clean_missing_and_needed_info
```


```{r}
write_csv(clean_missing_and_needed_info, "Data/my_exports/May13/fix_missing_urls.csv")
```


## timed out: copy paste discovery and fix

###### Okay let me check the event descriptions here

```{r}
count_all_highway_intox_anti_prep <- count_all_highway_intox %>%
  select(-n)

people_in_here_highway_intox_anti_prep <- people_in_here_highway_intox %>%
  select(-n)

highway_anti <- count_all_highway_intox_anti_prep %>%
  anti_join(people_in_here_highway_intox_anti_prep)  %>%
  inner_join(clean_scraped_prep) %>%
  #filter(is.na(reduced_offense_desc))
  group_by(offense_description, reduced_offense_desc) %>%
  tally() 
  #arrange(desc(n)) 

highway_anti
```

### Potentially more missing people

```{r}
clean_scraped_prep_missing <- clean_scraped_prep %>%
  filter(eventDescription == "Timeout error")

clean_scraped_prep_missing
```

```{r}
write_csv(clean_scraped_prep_missing, "code/Data/for_ryan_still_missing.csv")
```

